---
title: "Modelling approach for crafting environmental regulations under deep uncertainty: whale watching in Ojo de Liebre, Mexico"
author: "Emilio Rodríguez-Izquierdo"
date: "September 2018"
email: "13emil@gmail.com""
output:
  word_document: default
---

In this document you will find the scripts of the article figures.Bear in mind that you need to use the csv file with the Monte Carlo results (50,000 iterations)

***Scenario laissez-faire**

##Script of the results for whale abundance vs rate of habitat loss

Load the libraries
```{r librerias, eval=T}
library(plyr)
library(ggplot2)
library(data.table)
```

**Load the database and create a subset**

The full database is named "MC"" and the subset which only contains thevvalues of abundance and rate of habitat loss every 25 years, will be called "graph.dat".

```{r hojadatos, eval=T}
MC <- read.csv("C:/Users/emili/Desktop/Escenarios MonteCarlo SEPTIEMBRE2017/Nuevos Escenarios_Feb2018/ESC21_Hab_Kball.csv") ##note that you will have to use your own path 

MC$r_ballenas<-as.numeric(MC$r_ballenas)


graph.dat<- MC[,c(98,104,109,95,119,125,130,116,135)]  #these are the columns that correspond to the parameters we want to use

```

**Change from wide to long **

```{r mlt, eval=T}
graph.dat1 <- melt(setDT(graph.dat), measure.vars = patterns("PH", "P\\d"), 
                   value.name = c("PdH", "Pob"),
                   variable.name= "year")

# rename factor levels of variable (you need to have the package "forcats" installed)

graph.dat1[, year := forcats::lvls_revalue(year, c("25", "50", "75", "100"))][]
```


**Now we do the phase-diagram**
We first define the colours (color-blind friendly) for each year through the object "cbPalette". 
```{r graf1, eval=T}
#color-blind-friendly
cbPalette <- c("#999999", "#009E73", "#E69F00","#56B4E9")

ggplot(graph.dat1,aes(x=PdH, y=Pob, colour=year)) +
  geom_point(alpha =0.2) +
  geom_hline(aes(yintercept=1000),colour="black",linetype="twodash",size=.8)+
  
  scale_size_area(max_size=3,
                  breaks=c(0.025,0.032,max(graph.dat$r_ballenas))
  ) +
  
  scale_y_continuous(expand=c(0,0)) +
  scale_x_continuous (expand=c(0,0), breaks=c(20,40,60,80,100)) +
  expand_limits(y=c(0,3000),x=c(0,105)) +
  scale_colour_manual(values=cbPalette) +
  
  labs(x=expression("Rate of habitat loss (ha/year)"),
       y=expression("Whale abundance (individulas)")) +
  
  theme_classic() + 
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=18,face="bold"),
        axis.line.x = element_line(linetype="solid", colour="black"),
        axis.line.y = element_line(linetype="solid", colour="black"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background= element_rect(fill = "gray99"),
        panel.border=element_blank(),
        legend.position="none"
     ) 
```

##Script of the graph for the fitted curves every 25 years.

First, we load the package ggpmisc (Aphalo, 2016) that we used (through the "stat_poly_eq") to obtain the equations and the coefficients of determination. You need to save the image with at least 1650x740 dimension to avoid having the equations on top of the curves.

```{r ajuste, eval=T}
library(ggpmisc)

my.formula <- y ~ x

p25en25<-ggplot(graph.dat1,aes(x=PdH, y=Pob, colour=year)) +
  
  geom_smooth(formula = my.formula, size=1.6, se=FALSE) +
  
  stat_poly_eq(formula=my.formula,
               aes(label = paste(..eq.label.., ..rr.label.., sep = "*plain(\";\")~~")),
               parse = T)  +
  
  
  labs(x=expression("Rate of habitat loss (ha/year)"),
       y=expression("Whale abundance (individuals)")) +
  scale_colour_manual(values=cbPalette) +
  
  theme_classic() +
  
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=18,face="bold"),
        legend.position="none" ) +
  
  scale_y_continuous(expand=c(0,0)) +
  scale_x_continuous (expand=c(0,1), breaks=c(20,40,60,80,100)) 

p25en25
```


##Script of the figure 6 that relates the whale abundance and the number of boats.


**We do the graphs with the mean values every 5 years**

```{r graf_juntas, eval=T}
library(ggpubr)

pobl<-subset(MC[,c(93,103,94,96,97,98,99,100,101,
                   102,104,105,106,107,108,109,
                   110,111,112,113,95)])

promed_pobl<-apply(pobl,2,mean)
tiempo<-seq(0,100,by=5)

graf_promed_pobl<-data.frame(tiempo,promed_pobl)

graf_promed_pobl$tiempo<-as.factor(graf_promed_pobl$tiempo)

population<- ggplot(graf_promed_pobl, aes(x=tiempo, y=promed_pobl)) +
  
  geom_point(colour="black", size=3)  +
  
  geom_vline(aes(xintercept=as.numeric(tiempo[c(6)])),
             linetype=4, colour="black") +
  
  scale_y_continuous(expand=c(0,0),breaks=c(500,1000,1500)) +
  
  
  expand_limits(y=c(0,1700)) +
  
  
  labs(y=expression("Average number of whales")) +
  
  theme_classic() + 
  
  theme(axis.text=element_text(size=14),
        axis.title=element_text(size=20,face="bold"),
        axis.ticks.margin = unit(c(0,0,0,0), "lines"),
        axis.text.x=element_blank(),
        axis.title.x=element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linetype="dashed",colour="gray"),
        panel.grid.minor = element_blank(),
        panel.background= element_rect(fill = "gray99"),
        panel.border=element_blank(),
        legend.position="none"
  ) 



####################################################
#### Graph of the mean values of boats        ####

EMB <- read.csv("C:/Users/emili/Desktop/Escenarios MonteCarlo SEPTIEMBRE2017/Nuevos Escenarios_Feb2018/ESC21_Hab_Kball.csv")


EMB$r_ballenas<-as.numeric(MC$r_ballenas)


Emb_todo<- subset(MC[,c(23, 33, 24, 26, 27, 28, 29, 30, 31, 32,  34, 35, 36, 37, 38, 39, 40,
                        41, 42, 43, 25)])

tiempo<-seq(0,100,by=5)

promed_Emb<-apply(Emb_todo,2,mean)

graf_promed_emb<-data.frame(tiempo,promed_Emb)

graf_promed_emb$tiempo<-as.factor(graf_promed_emb$tiempo)


boats<- ggplot(graf_promed_emb, aes(x=tiempo, y=promed_Emb)) +
  
  geom_point(colour="black", size=3)  +
  geom_vline(aes(xintercept=as.numeric(tiempo[c(6)])),
             linetype=4, colour="black") +
  
  scale_y_continuous(breaks=c(20,40,60,80)) +
  scale_x_discrete(breaks=c(25,50,75,100)) +
  
  expand_limits(y=c(0,85)) +
  
  labs(x=expression("year"),
       y=expression("Average number of boats")) +
  
  theme_classic() + 
  
  theme(axis.text=element_text(size=14),
        axis.title=element_text(size=20,face="bold"),
        axis.ticks.margin = unit(c(0,0,0,0), "lines"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linetype="dashed",colour="gray"),
        panel.grid.minor = element_blank(),
        panel.background= element_rect(fill = "gray99"),
        panel.border=element_blank(),
        legend.position="none"
  ) 


figure<- ggarrange (population, boats, 
                    ncol=1, nrow=2, align = "v")

figure
```



***Scenario government intervention**



Load the libraries
```{r librerias, eval=T}
library(plyr)
library(ggplot2)
library(data.table)
```

**Load the database and create a subset**

The full database is named "MC"" and the subset which only contains thevvalues of abundance and rate of habitat loss every 25 years, will be called "graph.dat".

```{r hojadatos, eval=T}
MC <- read.csv("C:/Users/emili/Desktop/Ecological Modelling/Versión NUEVA_Junio2018/Resultados_30emb.csv") ##note that you will have to use your own path 

MC$r_ballenas<-as.numeric(MC$r_ballenas)


graph.dat<- MC[,c(98,104,109,95,119,125,130,116,135)]  #these are the columns that correspond to the parameters we want to use

```

**Change from wide to long **

```{r mlt, eval=T}
graph.dat1 <- melt(setDT(graph.dat), measure.vars = patterns("PH", "P\\d"), 
                   value.name = c("PdH", "Pob"),
                   variable.name= "year")

# rename factor levels of variable (you need to have the package "forcats" installed)

graph.dat1[, year := forcats::lvls_revalue(year, c("25", "50", "75", "100"))][]
```


**Now we do the phase-diagram**
We first define the colours (color-blind friendly) for each year through the object "cbPalette". 
```{r graf1, eval=T}
#color-blind-friendly
cbPalette <- c("#999999", "#009E73", "#E69F00","#56B4E9")

ggplot(graph.dat1,aes(x=PdH, y=Pob, colour=year)) +
  geom_point(alpha =0.2) +
  geom_hline(aes(yintercept=1000),colour="black",linetype="twodash",size=.8)+
  
  scale_size_area(max_size=3,
                  breaks=c(0.025,0.032,max(graph.dat$r_ballenas))
  ) +
  
  scale_y_continuous(expand=c(0,0)) +
  scale_x_continuous (expand=c(0,0), breaks=c(10,20,30)) +
  expand_limits(y=c(0,3000),x=c(0,35)) +
  scale_colour_manual(values=cbPalette) +
  
  labs(x=expression("Rate of habitat loss (ha/year)"),
       y=expression("Whale abundance (individulas)")) +
  
  theme_classic() + 
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=18,face="bold"),
        axis.line.x = element_line(linetype="solid", colour="black"),
        axis.line.y = element_line(linetype="solid", colour="black"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background= element_rect(fill = "gray99"),
        panel.border=element_blank(),
        legend.position="none"
     ) 
```

##Script of the graph for the fitted curves every 25 years.

First, we load the package ggpmisc (Aphalo, 2016) that we used (through the "stat_poly_eq") to obtain the equations and the coefficients of determination. You need to save the image with at least 1650x740 dimension to avoid having the equations on top of the curves.

```{r ajuste, eval=T}
library(ggpmisc)

my.formula <- y ~ x

p25en25<-ggplot(graph.dat1,aes(x=PdH, y=Pob, colour=year)) +
  
  geom_smooth(formula = my.formula, size=1.6, se=FALSE) +
  
  stat_poly_eq(formula=my.formula,
               aes(label = paste(..eq.label.., ..rr.label.., sep = "*plain(\";\")~~")),
               parse = T)  +
  
  
  labs(x=expression("Rate of habitat loss (ha/year)"),
       y=expression("Whale abundance (individuals)")) +
  scale_colour_manual(values=cbPalette) +
  
  theme_classic() +
  
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=18,face="bold"),
        legend.position="none" ) +
  
  scale_y_continuous(expand=c(0,0)) +
  scale_x_continuous (expand=c(0,1), breaks=c(10,20,30)) 

p25en25
```


##Script of the figure 6 that relates the whale abundance and the number of boats.


**We do the graphs with the mean values every 5 years**

```{r graf_juntas, eval=T}
library(ggpubr)

pobl<-subset(MC[,c(93,103,94,96,97,98,99,100,101,
                   102,104,105,106,107,108,109,
                   110,111,112,113,95)])

promed_pobl<-apply(pobl,2,mean)
tiempo<-seq(0,100,by=5)

graf_promed_pobl<-data.frame(tiempo,promed_pobl)

graf_promed_pobl$tiempo<-as.factor(graf_promed_pobl$tiempo)

population<- ggplot(graf_promed_pobl, aes(x=tiempo, y=promed_pobl)) +
  
  geom_point(colour="black", size=3)  +
  
  geom_vline(aes(xintercept=as.numeric(tiempo[c(6)])),
             linetype=4, colour="black") +
  
  scale_y_continuous(expand=c(0,0),breaks=c(500,1000,1500)) +
  
  
  expand_limits(y=c(0,1800)) +
  
  
  labs(y=expression("Average number of whales")) +
  
  theme_classic() + 
  
  theme(axis.text=element_text(size=14),
        axis.title=element_text(size=20,face="bold"),
        axis.ticks.margin = unit(c(0,0,0,0), "lines"),
        axis.text.x=element_blank(),
        axis.title.x=element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linetype="dashed",colour="gray"),
        panel.grid.minor = element_blank(),
        panel.background= element_rect(fill = "gray99"),
        panel.border=element_blank(),
        legend.position="none"
  ) 



####################################################
#### Graph of the mean values of boats        ####

EMB <- read.csv("C:/Users/emili/Desktop/Ecological Modelling/Versión NUEVA_Junio2018/Resultados_30emb.csv")


EMB$r_ballenas<-as.numeric(MC$r_ballenas)


Emb_todo<- subset(MC[,c(23, 33, 24, 26, 27, 28, 29, 30, 31, 32,  34, 35, 36, 37, 38, 39, 40,
                        41, 42, 43, 25)])

tiempo<-seq(0,100,by=5)

promed_Emb<-apply(Emb_todo,2,mean)

graf_promed_emb<-data.frame(tiempo,promed_Emb)

graf_promed_emb$tiempo<-as.factor(graf_promed_emb$tiempo)


boats<- ggplot(graf_promed_emb, aes(x=tiempo, y=promed_Emb)) +
  
  geom_point(colour="black", size=3)  +
  geom_vline(aes(xintercept=as.numeric(tiempo[c(6)])),
             linetype=4, colour="black") +
  
  scale_y_continuous(breaks=c(15,30)) +
  scale_x_discrete(breaks=c(25,50,75,100)) +
  
  expand_limits(y=c(0,40)) +
  
  labs(x=expression("year"),
       y=expression("Average number of boats")) +
  
  theme_classic() + 
  
  theme(axis.text=element_text(size=14),
        axis.title=element_text(size=20,face="bold"),
        axis.ticks.margin = unit(c(0,0,0,0), "lines"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linetype="dashed",colour="gray"),
        panel.grid.minor = element_blank(),
        panel.background= element_rect(fill = "gray99"),
        panel.border=element_blank(),
        legend.position="none"
  ) 


figure<- ggarrange (population, boats, 
                    ncol=1, nrow=2, align = "v")

figure
```

